Cores from kicks: constraints for observations of gravitational recoil in elliptical galaxies

### 2024-02-07 papers/paper-corekick/paper-corekick.tex
Writing on orbital stuff

@@

We observe a trend of the Voronoi bins with $h_4<0$ extending to larger radii for remnants with larger recoil kicks than for those remnants with smaller recoil kicks.
To quantify this recoil velocity dependence, we calculate a radially-varying mass-weighted $h_4$ parameter analogous to the observational spin parameter $\lambda$:
\begin{equation}
    \langle h_4 \rangle = \frac{\langle R h_4 \rangle}{\langle R \rangle},
\end{equation}
where the $\langle\rangle$ brackets indicate a mass-weighted average (thus equivalent to a flux weighted average in the case of constant mass-to-light ratio).
We observe an overall shift to more negative values of $h_4$ for kick velocities up until $\vk\sim400\,\kmps$ \drafting{check this, currently just read off plot}, above which the radial $h_4$ profile is not sensitive to increasing kick velocity. 
Additionally, all $h_4$ profiles are radially decreasing, obtaining a local minimum at $R\simeq 2\,\mathrm{kpc}$ ($R/\Reff\simeq 0.2$), and a modestly increasing for $R \gtrsim 2\,\mathrm{kpc}$.

@@

As seen in \autoref{fig:orbits}, we observe a clear trend of higher kick velocity remnants having a higher fraction of $\pi$-box orbits, and a lower fraction of $x$-tube orbits, than their lower kick velocity counterparts.
The other five orbital families are generally consistent between all merger remnants. 
\drafting{
    Might move this to discussion, just put here for now to get the idea down.
    z-tubes not primary contributor to LOS velocity (rotation about z axis).
    Higher kick correlated with reduction in x-tubes (orbits about the x-axis), increase in pi-box.
    Pi-box show no resonant motion (no net ang. mom.)\dots here's the leap: reduced resonant motion seen as a reduction in `symmetric deviations' in line profile?
    Also look directly at kinematic maps: 0 kick case. Dominated by x tubes at radii > 1kpc, small region of negative h4 extends to here before becoming positive. 
}

>>>>

Following \citet{nasim2021b}, we fit a power law of the form
\begin{equation}\label{eq:vkrb}
    \frac{\rb}{r_{\mathrm{b},0}} = K \left( \frac{\vk}{v_\mathrm{esc}} \right)^\beta + 1
\end{equation}
to our simulations, which is thus valid for $0\leq\vk/\kmps\leq900$, or equivalently $0\leq\vk/v_\mathrm{esc}\leq 0.75$.
\drafting{We estimate the parameters $K$ and $\beta$ using HMC, and find median values of $K=$ and $\beta=$.}

\drafting{Using \autoref{eq:vkrb}, we push through the distribution of kick velocities from \citet{zlochower2015} for dry mergers assuming random azimuthal alignment (their Fig. 15, right panel.) 
This allows us to obtain a distribution of core radius values predicted by a given kick velocity model.
We find\dots}

@@

We observe a trend of the Voronoi bins with $h_4<0$ extending to larger radii for remnants with larger recoil kicks than for those remnants with smaller recoil kicks.
To quantify this recoil velocity dependence, we calculate a radially-varying mass-weighted $h_4$ parameter analogous to the observational spin parameter $\lambda$:
\begin{equation}
    \langle h_4(R') \rangle = \frac{\langle R h_4 \rangle}{\langle R \rangle} = \frac{\sum_i^N M_i h_{4,i} R_i}{\sum_i^N M_i R_i} \;\forall R_i \leq R',
\end{equation}
where the $\langle\rangle$ brackets indicate a mass-weighted average (thus equivalent to a flux weighted average in the case of constant mass-to-light ratio).
We observe an overall shift to more negative values of $h_4$ for kick velocities up until $\vk\sim400\,\kmps$ \drafting{check this, currently just read off plot}, above which the radial $h_4$ profile is not sensitive to increasing kick velocity. 
Additionally, all $h_4$ profiles are radially decreasing, obtaining a local minimum at $R\simeq 2\,\mathrm{kpc}$ ($R/\Reff\simeq 0.2$), and a modestly increasing for $R \gtrsim 2\,\mathrm{kpc}$.

@@

As seen in \autoref{fig:orbits}, we observe a clear trend of higher kick velocity remnants having a higher fraction of $\pi$-box orbits, and a lower fraction of $x$-tube orbits, than their lower kick velocity counterparts.
The other five orbital families are generally consistent between all merger remnants. 
As discussed in \citet{frigo2021}, negative values of $h_4$ indicate regions with weaker tails of the LOS velocity distribution due to the superposition of radial orbits with different orientations.
This explanation supports our observation of an increase in $\pi$-box orbits (which have no net angular momentum are typically very radial orbits) with kick velocity, and a corresponding decrease in the value of $h_4$ (comparing the IFU maps in \autoref{fig:IFU0} to \autoref{fig:IFU900}).
Equivalently, $x$-tube orbits (which rotate about the major axis of the inertia tensor) appear as a peaked LOS velocity distribution in the projection we use, leading to extended tails, and hence a more positive value of $h_4$ than for a weak tail distribution. 
All together, we can infer that as a SMBH ploughs through the surrounding stellar environment, regular $x$-tube orbits are disrupted to non-resonant $\pi$-box orbits.
With a higher kick velocity, the SMBH is able to cumulatively disturb orbits at a larger radial extent, converting a greater fraction of $x$-tube orbits to $\pi$-box orbits, resulting in the increased expanses of negative $h_4$ regions in the kinematic maps.
This also explains the overall downward shift in the $\langle h_4 \rangle$ profiles in \autoref{fig:h4} for higher kick velocities. 
\drafting{Why are the $\langle h_4 \rangle$ profiles consistent in shape though??}

\drafting{
    Might move this to discussion, just put here for now to get the idea down.
    z-tubes not primary contributor to LOS velocity (rotation about z axis).
    Higher kick correlated with reduction in x-tubes (orbits about the x-axis), increase in pi-box.
    Pi-box show no resonant motion (no net ang. mom.)\dots here's the leap: reduced resonant motion seen as a reduction in `symmetric deviations' in line profile?
    Also look directly at kinematic maps: 0 kick case. Dominated by x tubes at radii > 1kpc, small region of negative h4 extends to here before becoming positive. 
}
